\LARGE \bf Enhancing Security in Multi-Robot Systems through Co-Observation Planning, Reachability Analysis, and Network Flow

### 2024-01-19 paper.tex
finished version

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8
\documentclass[journal]{IEEEtran}  % Comment this line out

\IEEEoverridecommandlockouts                       % This command is only
% needed if you want to
% use the \thanks command
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org

@@

\usepackage{cleveref}
\graphicspath{{Figs/}}

@@

Enhancing Security in Multi-Robot Systems through Co-Observation Planning on Unsecured Trajectories}

\author{Ziqi Yang and Roberto Tron\thanks{This project is supported by the National Science Foundation grant "CPS: Medium: Collaborative Research: Multiagent Physical Cognition and Control Synthesis Against Cyber Attacks" (Award number 1932162).}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
This paper addresses the critical issue of security in multi-robot systems (MRS) when facing threats from adversaries attempting to compromise the robots' control which could lead to unauthorized access to forbidden areas, potentially causing harm or disrupting the mission. To tackle this problem, we propose a novel multi-robot planning algorithm that leverages mutual observations between robots, and introducing  \emph{reachability} constraints as additional security measures. By enforcing an empty intersection between the locations that the agents could potentially reach (using reachability constraint) and the forbidden regions, we provide guarantees that even in the face of adversarial movement by attackers, compromised robots cannot reach forbidden regions without missing scheduled co-observations. The reachability constraint is implemented with ellipsoidal over-approximation; this allows to quickly check intersection and compute gradients.

Furthermore, to further enhance system resilience and address the feasibility challenges encountered in the initial stage, we introduce the concept of \emph{sub-teams}, where multiple robots form cohesive units along each route, replacing individual robot assignments. This allows redundant robots to deviate for co-observations across different sub-teams, ensuring the security of multiple trajectories without the need for trajectory modifications of the entire sub-team. To plan cross-trajectory co-observations that maintain security against plan-deviation attacks, we formulate the multi-flow problem on unsecured MRS trajectories,

We demonstrate the effectiveness and robustness of our proposed algorithm, which significantly strengthens the security of multi-robot systems in the face of adversarial threats.
\end{abstract}

\section{Introduction}\label{sec:introduction}
Multi-robot systems (MRS) have found wide applications in various fields, including warehouse organization, surveillance and precision agriculture, owing to their capacity for complex task execution through cooperation and coordination \cite{pajares2015overview, julian2012distributed}. While MRS offers numerous advantages, their distributed nature and dependence on network communication render them vulnerable to cyber threats, introducing new cybersecurity challenges such as unauthorized access, malicious attacks, and data manipulation \cite{brunner2010infiltrating}. This paper addresses the specific scenario in which attackers compromise robots and direct them to \emph{forbidden regions}, which may contain security-sensitive equipment or human workers. Such countering deliberate deviations, termed \emph{plan-deviation attacks}, are identified and addressed in previous studies \cite{wardega2019resilience, wardega2023byzantine, wardega2023hola, yang2021multi, yang2020multi}. As a security measure, robots perform inter-robot co-observations using their onboard sensing capabilities to check for unusual behavior. These mutual observation establish an co-observation schedule alongside the path, ensuring that any attempts by a compromised robot to violate safety constraints (such as transgressing forbidden regions) would break the observation plan and be promptly detected.

A preliminary version of the paper was presented at conferences \cite{yang2020multi,yang2021multi}. When extanding the solution from \cite{wardega2019resilience} in grid-world to continuous configuration spaces, additional reachability analysis is erquired for additional security. Reachability analysis is a important step in the security and safety verification synthesis for cyber-physical systems (CPSs) \cite{gueguen2009safety,ding2020secure}. One typical approach involves computing an over-approximation of the reachable space for checking safety properities. Previous studies \cite{kurzhanski2000ellipsoidal,lakhal2019interval,maiga2015comprehensive} often use geometric extrapolations, such as zonotopes and ellipsoids, to represent a compact enclosure of the reachability sets. In the realm of online safety assessment against cyber attacks, \cite{kwon2017reachability} computed the reachable set of CPS states achievable by potential cyber attacks and compared it with the safe region, set based on current state estimation and environmental conditions. In contrast, our work shifts the focus to reachability analysis during the planning phase, addressing the path optimization problem by incorporating constraints based on sets of locations that agents could potentially reach, referred to as \emph{reachability regions}. This contribution adds a novel perspective to the existing literature.

This proposed approach enforces an empty intersection between forbidden regions and an ellipsoidal reachability region during trajectory optimization, preventing undetected attacks if an adversary gains control of the robots. Inspired by the \emph{heuristic sampling domain} introduced by \cite{gammell2014informed} within the context of the $\mathtt{RRT^*}$ path planning algorithm, we utilize an ellipsoidal boundary to constrain the search space and formulate the ellipsoid as the reachability region constraint. We present a mathematical formulation of reachability regions compatible with the solver as a spatio-temporal constraint.

This paper also extend the previous works by fixing two main challenges identified in \cite{wardega2023hola, yang2020multi, yang2021multi}, the feasibility problem and optimality problem. Firstly, finding a co-observation and reachability secured plan may not always be feasible when robots are separated from others by obstacles or forbidden regions, or are located far away from each other. Secondly, security requirements may come at the cost of overall system performance, which is particularly concerning given that system performance is a key factor in the decision to use multi-agent systems. In \Cref{sec:cross-trajectory}, these challenges are addressed through introducing additional robots into the MRS. Robots are formed into \emph{sub-teams} and tasked to follow an optimal trajectory without security constraints. The added redundant robots are then used to setup additional co-observations both within the sub-team and across different sub-teams by deviating from the . Cross-trajectory co-observations allows sub-teams to preserve the optimal unsecured trajectories (as shown in Figure \ref{fig:cross-traj-comparison-set}), as it does not require the entire \emph{sub-team} to maintain a close distance with other teams when co-observation occurs.

@@

\paragraph*{Paper contributions} 
In this paper, two main contributions have been presented.

@@

\paragraph*{paper ourline}
The rest of this paper is organized as follows. In \Cref{sec:ADMM-planning}, we begin by providing an overview of the ADMM-based optimal trajectory solver. We then introduced the ellipsoidal reachability region and formulate it to different constraints with respect to different type of forbidden regions. In \Cref{sec:cross-trajectory}, we proposed the formulation of sub-teams and introduced the planning algorithm of the cross-trajectory co-observation planning problem. Finally, in \Cref{sec:summary}, we provide summary of this paper.

\section{Literature review}
\subsection{Reachability analysis}
Reachability analysis is a common step in the security and safety verification synthesis for cyber-physical systems (CPSs) \cite{gueguen2009safety,ding2020secure}. A common method is to compute an over-approximation of the reachable space for checking safety properities. Various researches rely on a geometrical extrapolation of the system states like zonotopes and ellipsoids as representation of a compact enclosure of the reachability sets \cite{kurzhanski2000ellipsoidal,lakhal2019interval,maiga2015comprehensive,maiga2015comprehensive}. The work in \cite{}

@@

In this section, we aim to enhance the security of MRSs against malicious takeovers and deviations by attackers. Specifically, we design an inter-robot observation plan (co-observation schedule) to ensures that, during the task period, robots are constantly in proximity to one another according to a schedule, in order to observe and detect potential hazardous behaviors. 
This is achieved by keeping the potential region that the robots can possibly reach between each consecutive co-observations away from the forbidden regions. 
\begin{definition}\label{def:secured-plan}
  A multi-robot trajectory plan is secure again plan-deviation attacks if it ensure that any potential deviations to these forbidden regions will cause the corresponding robot to miss their next co-observation with other robots. 
\end{definition}

This problem is first solved by \cite{wardega2019resilience} in grid world for robots with fixed start and goal locations (Figure~\ref{fig:Grid-example-application}) but does not offer the ability to optimize arbitrary smooth cost functions of the trajectories.

We formulate the planning problem as an optimal trajectory optimization problem to minimize arbitrary smooth objective functions. 
We denote as $q_{ij}\in\real{m}$ the position of agent $i$ at the discrete-time index $j$, with $m$ representing the dimension of the state space. For a total of $n_p$ robots, and a task time horizon of $T$, a total of $n_p$ trajectories can be represented as an aggregated vector $\vq\in \real{n m T}$. 
The goal of our path planning problem is to minimize or maximize an objective function $\varPhi(\vq)$ under a set of nonlinear constraints described by a set $\Omega$. The set $\Omega$ is given by the intersection of spatio-temporal sets given by  security constraints (co-observation schedule, reachability analysis) and traditional path planning constraints. Formally:
\begin{equation}\label{eq:general-problem}
	\begin{split}
		\min/\max & \quad \varPhi(\vq)\\
		\textrm{subject to} &\quad \vq \in \Omega.
	\end{split}
\end{equation}

@@

\begin{example} 
We consider the estimation of a slowly-varying scalar or vector field, denoted as $\vx$, at discrete locations within an unknown environment. Autonomous agents navigate this environment, collecting sensory data to construct a corresponding map (see Figure~\ref{fig:ADMM-example-application}). Our map comprises points of interest arranged on a grid, each associated with a slowly-changing value. Our goal is to find paths that minimize uncertainty and effectively reconstruct the field. 
We associate a Kalman Filter (KF)~(cf. \cite{anderson2012optimal}) to each point $j$, and use it to track the uncertainty through its estimated covariance $P_j$. The updates of the filters are based on measurements from the robot centered at $q$, and we use a Gaussian radial basis function  modeling spatially-varying measurement quality.

\rtron{Cite conference article for details on the problem formulation.}
The optimization objective $\varPhi(\vq)$ is the maximum uncertainty $\max_j P_{j}$. The set $\Omega$ includes traditional constraints like bounded velocity, convex obstacles, waypoints to reach with deadlines, in addition to security constraints including co-observation and reachability constraints.
\end{example}

\begin{figure}
     \centering
     \subfloat[Continuous world trajectory \label{fig:ADMM-example-application}]{\includegraphics[width=0.45\linewidth, trim = 2cm 2cm 2cm 2cm]{Example_application}}
     \subfloat[Grid-world trajectory. \label{fig:Grid-example-application}]{\includegraphics[width=0.45\linewidth, trim = 2cm 0.5cm 0cm 2cm]{Kacper_result}}
     \caption{Trajectory design of an example map exploration task for a three-robot system with start locations and destinations fixed. Task are planned in a $8 \times 8$ grid world. Zone 1 is obstacle, Zone 2 and Zone 3 are safe zones. Fig~\ref{fig:Grid-example-application} shows the planned trajectory with co-observation schedule in grid-world. Fig~\ref{fig:ADMM-example-application} shows the trajectory design in continuous world optimized with respect to a map exploration task}
     \label{fig:example-application}
 \end{figure}
 \rtron{Make sure we refer back to Example 1 when makes sense}

 This section is organized as follows. We begin with foundational mathematical concepts, such as the basic ADMM optimal trajectory solver. Subsequently, our solver allows us to seamlessly integrate security constraints, encompassing co-observation schedule and reachability constraints. Concluding this section, we present the obtained results alongside limitations (due to infeasibility) that are then addressed in \Cref{sec:cross-trajectory}.

\subsection{Preliminaries}
We review in this section the mathematical foundations for our ADMM solver and the constraints that we will use.

\subsubsection{Differentials}
We define the differential of a map $f(x):\real{m}\to\real{n}$ at a point $x_0$ as the unique matrix $\partial_x f \in\real{n\times m}$ such that
\begin{equation}\label{equ:dt_to_dx}
  \left.\dert f\bigl(x(t)\bigr)\right|_{t=0}=\partial_x f\bigl(x(0)\bigr) \dot{x}(0)
\end{equation}
where $t\mapsto x(t)\in\real{n}$ is a smooth parametric curve such that $x(0)=x$ with any arbitrary tangent $\dot{x}(0)$. For brevity we will use $\dot f$ for $\dert f$ and $\partial_x f$ for $\frac{\partial f}{\partial x}$. 
%The differentials $\partial_x f$ is derived through \eqref{equ:dt_to_dx} having $\dot f$ divided by $\dot x$.

@@

 \subsubsection{Householder rotations}\label{sec:householder}
 In the derivation of the reachability constraint in \Cref{sec:reachability,sec:ellipse-region-constraint}, we need to define a differentiable transformation to a canonical ellipse. This transformation includes a rotation derived from a modified version of Householder transformations \cite{householder1958unitary}. With respect to the standard definition, our modification ensures that the final operator is a proper rotation (i.e., not a reflection). We call our version of the operator a \emph{Householder rotation}. In this section we derive Householder rotations and their differentials for the 3-D case; the 2-D case can be easily obtained by embedding it in the $z=0$ plane.

  \begin{definition} Let $\nu_\cF$ and $\nu_\cE$ be two unitary vectors ($\norm{\nu_\cF}=\norm{\nu_\cE}=1$). Define the normalized vector $u$ as
    \begin{equation}
      u'=\nu_\cF+\nu_\cE,\quad u =\frac{u'}{\norm{u'}}.
    \end{equation}
    The \emph{Householder rotation} $H(\nu_\cF,\nu_\cE)$ is defined as
    \begin{equation}\label{eq:H definition}
      H(\nu_\cF,\nu_\cE) = 2 u u\transpose-I.
    \end{equation}
  \end{definition}
  The main property of interest for our application is the fact that $H$ is a rotation mapping $\hat{\nu_\cF}$ to $\hat{\nu_\cE}$, as shown by the following.
  \begin{proposition}\label{prop:HProperty}
    The matrix $H$ has the following properties:
    \begin{enumerate}
    \item It is a rotation, i.e.
      \begin{enumerate}
      \item\label{it:orthonormality} $H\transpose H=I$;
      \item\label{it:determinant} $\det(H)=1$.
      \end{enumerate}
    \item\label{it:transformation} $\nu_\cE=H \nu_\cF$.
    \end{enumerate}
  \end{proposition}
  \begin{proof}
  	See Appendix \ref{proof:HProperty}.
  \end{proof}

@@

\subsubsection{Alternating Directions Method of Multipliers (ADMM)}\label{chapter:ADMM review}
The basic idea behind our ADMM-based solver \cite{yang2020multi} is to split the constraints from the objective function using a different set of variables $\vz$, and then solve an Augmented Lagrangian formulation of the optimization problem in \eqref{eq:general-problem}.
More specifically, we can rewrite the constraint $\vq\in\varOmega$ using an indicator function $\varTheta$, and include it in the objective function. 
In the traditional application of ADMM, the variables $\vz$ are simple duplicates of $\vx$ projected to the constraint set $\Omega$. However, in path planning problems, some constraints are non-convex, rendering the projection step more difficult (due to the presence of multiple local minima).
To allow for an easier projection step, we propose a minor generalization of the ADMM formulation where we allow $\vz$ to replicate an arbitrary function of the main variables $\vq$ (instead of being an exact copy), to transform constraint $\vq\in\Omega$ to $D(\vq) \in \sZ$. In summary, we transform \Cref{eq:general-problem} into
\begin{equation}\label{eq:ADMMSetConstraint_modified}
	\begin{aligned}
		&\max\quad \varPhi(\vq)+\varTheta(\vz) \\
		& \begin{array}{r@{\quad}c}
			s.t.& D(\vq)-\vz=0
		\end{array} 
	\end{aligned}
\end{equation}
where $D(\vq)= [D_1(\vq)^T,\dots,D_l(\vq)^T]^T$ is a vertical concatenation of different functions for different constraints.

The update steps of the ADMM algorithm are then derived as \cite{Boyd2011}:
\begin{subequations}\label{eq:ADMMupdate}
	\begin{align}
		\label{eq:admm-mod-update-q}\vq^{k+1}&:=\argmin_\vq(\varPhi(\vq^k) +\frac{\rho}{2}\norm{D(\vq)-\vz^k+\vu^k}_2^2),\\
		\vz^{k+1}&:=\Pi_\mathcal{Z}(D(\vq^{k+1})+\vu^k), \\
		\vu^{k+1}&:=\vu^k+D(\vq^{k+1})-\vz^{k+1},
	\end{align}
\end{subequations}
where $\Pi_\mathcal{Z}$ is the new projection to the modified constraint set $\mathcal{Z}$, $\vu$ represents a scaled dual variable that, intuitively, accumulates the sum of primal residuals
\begin{equation}\label{eq:primal-residual}
	\vr^{k}=D(\vq^{k+1})-\vz^{k+1}.
\end{equation}
Checking the primal residuals alongside with the dual residuals 
\begin{equation}\label{eq:dual-residual}
	\vs^{k}=-\rho(\vz^{k}-\vz^{k-1})
\end{equation}
after each iteration, the steps are reiterated until convergence when the primal and dual residuals are small, or divergence when primal and dual residual remains large after a fixed number of iterations.
Since $D(\vq)$ is the vertical concatenation of $D_i(\vq)$, the projection of each set $\mathcal{Z}_i$ is independent for each constraint and can be computed separately. The advantage of this formulation is that we can choose $D(\vq)$ such that the new constraint set~$\cZ$ becomes simple to compute, and the function~$D(\vq)$ can be used to select only the subset of the variables on which a constraint depends, speeding up computations; however, the drawback is that we move the non-convexity to the primal cost function, i.e., in the update for $\vq$ in \eqref{eq:admm-mod-update-q}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8
%\documentclass[journal]{IEEEtran}  % Comment this line out
\documentclass[10pt,twocolumn,twoside]{IEEEtran} 
\IEEEoverridecommandlockouts                       % This command is only
% needed if you want to
% use the \thanks command
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

%\documentclass[journal,twoside,web]{ieeecolor}
% The following packages can be found on http:\\www.ctan.org

@@

\usepackage[export]{adjustbox} 
\usepackage{cleveref}
\graphicspath{{Figs/}}

@@

Enhancing Security in Multi-Robot Systems through Co-Observation Planning, Reachability Analysis, and Network Flow}

\author{Ziqi Yang, Roberto Tron \IEEEmembership{Member, IEEE} 
\thanks{This project is supported by the National Science Foundation grant "CPS: Medium: Collaborative Research: Multiagent Physical Cognition and Control Synthesis Against Cyber Attacks" (Award number 1932162).}
\thanks{Ziqi Yang is with the Department of Systems Engineering,
Boston University, Boston, MA 02215 USA (e-mail: zy259@bu.edu).
}
\thanks{Roberto Tron is with the Faculty of Mechanical Engineering and Systems Engineering, Boston University, Boston, MA 02215 USA (e-mail:
tron@bu.edu).}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
This paper addresses security challenges in multi-robot systems (MRS) where adversaries may compromise robot control, risking unauthorized access to forbidden areas. We propose a novel multi-robot optimal planning algorithm that integrates mutual observations and introduces \emph{reachability} constraints for enhanced security. This ensures that, even with adversarial movements, compromised robots cannot breach forbidden regions without missing scheduled co-observations. The reachability constraint uses ellipsoidal over-approximation for efficient intersection checking and gradient computation. To enhance system resilience and tackle feasibility challenges, we also introduce \emph{sub-teams}. These cohesive units replace individual robot assignments along each route, enabling redundant robots to deviate for co-observations across sub-teams, securing multiple trajectories without requiring modifications. The planning of cross-trajectory co-observations that maintain security against plan-deviation attacks is sloved by formulating the multi-flow problem on unsecured MRS trajectories. We demonstrate the effectiveness and robustness of our proposed algorithm, which significantly strengthens the security of multi-robot systems in the face of adversarial threats.
\end{abstract}

\section{Introduction}\label{sec:introduction}
Multi-robot systems (MRS) have found wide applications in various fields. While offering numerous advantages, the distributed nature and dependence on network communication render the MRS vulnerable to cyber threats, such as unauthorized access, malicious attacks, and data manipulation \cite{brunner2010infiltrating}. This paper addresses a specific scenario in which  robots are compromised by attacker and directed to \emph{forbidden regions}, which may contain security-sensitive equipment or human workers. Such countering deliberate deviations, termed \emph{plan-deviation attacks}, are identified and addressed in previous studies \cite{wardega2019resilience, wardega2023byzantine, wardega2023hola, yang2021multi, yang2020multi}. As a security measure, we utilized onboard sensing capabilities of the robots to perform inter-robot co-observations and check for unusual behavior. These mutual observation establish a co-observation schedule alongside the path, ensuring that any attempts by a compromised robot to violate safety constraints (such as transgressing forbidden regions) would break the observation plan and be promptly detected.

A preliminary version of the paper was presented in \cite{yang2020multi,yang2021multi}. Extanding the grid-world solution from \cite{wardega2019resilience} to continuous configuration spaces, we incorporate the co-observation planning as constraints in the ADMM-based trajectory optimization solver to accomodate for more flexible objectives. In this paper, we incorporate additional reachability analysis during the planning phase, incorporating constraints based on sets of locations that agents could potentially reach, referred to as \emph{reachability regions}, as a novel perspective to the existing literature. This approach enforces an empty intersection between forbidden regions and the reachability region during trajectory optimization, preventing undetected attacks if an adversary gains control of the robots. Inspired by the \emph{heuristic sampling domain} introduced by \cite{gammell2014informed} within the context of the $\mathtt{RRT^*}$ path planning algorithm, we utilize an ellipsoidal boundary to constrain the search space and formulate the ellipsoid as the reachability region constraint. We present a mathematical formulation of reachability regions compatible with the solver in \cite{yang2020multi} as a spatio-temporal constraint.

However, such plans are not guaranteed to exist, and the secured trajectory always come at the cost of overall system performance. In this paper, we also address these feasibility and optimality challenges identified in \cite{wardega2023hola, yang2020multi, yang2021multi}. For a MRS with unsecured optimal trajectorys (without security constraints), we introduce redundant robots and formed into \emph{sub-teams} with the original ones. These redundant robots are assigned the task of establishing additional co-observations, termed \emph{cross-trajectory co-observations}, within and across different sub-teams. The proposed algorithm focuses on the movement plans for the additional robots, distinct from those dedicated to task objectives, indicating when they should stay with their current sub-team and when they should deviate to join another. This strategy allows sub-teams to preserve the optimal unsecured trajectories (as illustrated in Figure \ref{fig:cross-traj-comparison-set}) without requiring the entire \emph{sub-team} to maintain close proximity to other teams during co-observation events. The cross-trajectory co-observation problem is transformed as a multi-agent path finding problem on roadmap (represented as directed graph) and solved as a network flow problem \cite{yu2013multi}.

@@

\noindent\myparagraph{Related research} Trajectory planning problem of MRS systems remained as a subject of intense study for many decades, and optimization is a common approaches in such area. Optimization based approaches are customizable to a variable of constraints (e.g. speed limit, avoid obstacles) and task specifications (e.g. maximum surveillance coverage, minimal energy cost). In constract to our work in \Cref{sec:ADMM-planning}, many motion planning tasks require a non-convex constraint problem formulation, most contributors focused on convex problems, and only allow for a few types of pre-specified non-convex constraints through convexification\cite{liu2014solving}\cite{VanParys2016}\cite{Schulman2014}. Several optimization techniques like MIQP~\cite{mellinger2012mixed} and ADMM~\cite{bento2013message} have been used to reduce computational complexity, and to incorporate more complex non-convex constraints.

Graph-based search is another extensively explored approach in trajectory planning problems, such as formation control \cite{tanner2004leader,hu2019distributed} and Multi-Agent Path Finding (MAPF) problems \cite{stern2019multi}. In traditional MAPF formulations, environments are abstracted as graphs, with nodes representing positions and edges denoting possible transitions between positions and solved as a network flow problem \cite{yu2013multi,yu2016optimal}. This formulation allows for the application of combinatorial network flow algorithms and linear program techniques, offering efficient and more flexible solutions to the planning problem. %In contrast with such approach, our research in \Cref{sec:cross-trajectory} abstracts the existing Multi-Robot System (MRS) trajectory (instead of the environment) into a directed graph and formulates the cross-trajectory co-observation paths as flows.

In \Cref{sec:ADMM-planning}, we incorporate reachability constraints to increase overall security Reachability analysis is a important step in the security and safety verification synthesis for cyber-physical systems (CPSs) \cite{gueguen2009safety,ding2020secure}. One typical approach involves computing an over-approximation of the reachable space for checking safety properities. Previous studies \cite{kurzhanski2000ellipsoidal,lakhal2019interval,maiga2015comprehensive} often use geometric representation, such as zonotopes and ellipsoids, as a compact enclosure of the reachability sets. In the realm of online safety assessment against cyber attacks, \cite{kwon2017reachability} computed the reachable set of CPS states achievable by potential cyber attacks and compared it with the safe region, set based on current state estimation and environmental conditions.

\noindent\myparagraph{Paper contributions} 
In this paper, two main contributions have been presented.

@@

\noindent\myparagraph{Paper Outline}
The rest of this paper is organized as follows. In \Cref{sec:ADMM-planning}, we begin by providing an overview of the ADMM-based optimal trajectory solver. We then introduced the ellipsoidal reachability region and formulate it to different constraints with respect to different type of forbidden regions. In \Cref{sec:cross-trajectory}, we proposed the formulation of sub-teams and introduced the planning algorithm of the cross-trajectory co-observation planning problem. Finally, in \Cref{sec:summary}, we provide summary of this paper.

@@

In this section, we design an inter-robot observation plan (co-observation schedule) to ensures that, during the task period, robots are constantly in proximity to one another according to a schedule, in order to observe and detect potential hazardous behaviors. 
This is achieved by keeping the potential region that the robots can possibly reach between each consecutive co-observations away from the forbidden regions. 
\begin{definition}\label{def:secured-plan}
  A multi-robot trajectory plan is secure again plan-deviation attacks if it ensure that any potential deviations to these forbidden regions will cause the corresponding robot to miss their next co-observation with other robots. 
\end{definition}

%This problem is first solved by \cite{wardega2019resilience} in grid world for robots with fixed start and goal locations (Figure~\ref{fig:Grid-example-application}) but does not offer the ability to optimize arbitrary smooth cost functions of the trajectories.

We formulate the planning problem as an optimal trajectory optimization problem to minimize arbitrary smooth objective functions. We denote as $q_{ij}\in\real{m}$ the position of agent $i$ at the discrete-time index $j$, with $m$ representing the dimension of the state space. For a total of $n_p$ robots, and a task time horizon of $T$, a total of $n_p$ trajectories can be represented as an aggregated vector $\vq\in \real{n m T}$. 
The overall goal is to minimize or maximize an objective function $\varPhi(\vq)$ under a set of nonlinear constraints described by a set $\Omega$, which is given by the intersection of spatio-temporal sets given by traditional path planning constraintsand the security constraints (co-observation schedule, reachability analysis). Formally:
\begin{equation}\label{eq:general-problem}
	\begin{split}
		\min/\max & \quad \varPhi(\vq)\\
		\textrm{subject to} &\quad \vq \in \Omega.
	\end{split}
\end{equation}

@@

\begin{example} 
Robots are tasked to navigate, collecting sensory data to construct a corresponding map for the slowly-varying field, denoted as $\vx$, of an unknown environment (see Figure~\ref{fig:SecurityBreak}). The map comprises points of interest arranged on a grid, each associated with a slowly-changing value. Our goal is to find paths that minimize uncertainty and effectively reconstruct the field. 
We associate a Kalman Filter (KF)~(cf. \cite{anderson2012optimal}) to each point $j$, and use it to track the uncertainty through its estimated covariance $P_j$. The updates of the filters are based on measurements from the robot centered at $q$, and we use a Gaussian radial basis function  modeling spatially-varying measurement quality. The optimization objective $\varPhi(\vq)$ is written as the maximum uncertainty $\max_j P_{j}$ (detailed formulation can be found in \cite{yang2020multi}). %The set $\Omega$ includes traditional constraints like bounded velocity, convex obstacles, waypoints to reach with deadlines, in addition to security constraints including co-observation and reachability constraints.
\end{example}

\begin{figure}
     \centering
     \subfloat[Grid-world trajectory. \label{fig:Grid-example-application}]{\includegraphics[width=0.45\linewidth, trim = 2cm 0.5cm 0cm 2cm]{Kacper_result}}
     \subfloat[Continuous world trajectory \label{fig:SecurityBreak}]{\includegraphics[width=0.45\linewidth, trim = 2cm 2cm 2cm 2cm]{SecurityBreak}}
     \caption{Trajectory design of an example map exploration task for a three-robot system with start locations and destinations fixed. Task are planned in a $8 \times 8$ grid world. Zone 1 is obstacle, Zone 2 and Zone 3 are safe zones. Fig~\ref{fig:Grid-example-application} shows the planned trajectory with co-observation schedule in grid-world. Fig~\ref{fig:SecurityBreak} shows the unsecured trajectory design in continuous world optimized with respect to a map exploration task where Zone 1 and 2 are considered together as one obstacle.  A potential path that could break the security by entering region 1 is highlighted in dashline.}
     \label{fig:example-application}
 \end{figure}

% This section is organized as follows. We begin with foundational mathematical concepts needed for the optimal trajectory solver and reachability constraints. Subsequently, we integrate security constraints, including co-observation schedule and reachability constraints. Finally, we present the obtained results alongside limitations that are then addressed in \Cref{sec:cross-trajectory}.

\subsection{Preliminaries}
\subsubsection{Differentials}
We define the differential of a map $f(x):\real{m}\to\real{n}$ at a point $x_0$ as the unique matrix $\partial_x f \in\real{n\times m}$ such that
\begin{equation}\label{equ:dt_to_dx}
  \left.\dert f\bigl(x(t)\bigr)\right|_{t=0}=\partial_x f\bigl(x(0)\bigr) \dot{x}(0)
\end{equation}
where $t\mapsto x(t)\in\real{n}$ is a smooth parametric curve such that $x(0)=x$ with any arbitrary tangent $\dot{x}(0)$. For brevity we will use $\dot f$ for $\dert f$ and $\partial_x f$ for $\frac{\partial f}{\partial x}$. 
%The differentials $\partial_x f$ is derived through \eqref{equ:dt_to_dx} having $\dot f$ divided by $\dot x$.

@@

 \subsubsection{Householder rotations}\label{sec:householder}
  We define a differentiable transformation to a canonical ellipse that is used in the derivation of the reachability constraints in \Cref{sec:reachability,sec:ellipse-region-constraint}. This transformation includes a rotation derived from a modified version of Householder transformations \cite{householder1958unitary}. With respect to the standard definition, our modification ensures that the final operator is a proper rotation (i.e., not a reflection). We call our version of the operator a \emph{Householder rotation}. In this section we derive Householder rotations and their differentials for the 3-D case; the 2-D case can be easily obtained by embedding it in the $z=0$ plane.

  \begin{definition} Let $\nu_\cF$ and $\nu_\cE$ be two unitary vectors ($\norm{\nu_\cF}=\norm{\nu_\cE}=1$). Define the normalized vector $u$ as
    \begin{equation}
      u'=\nu_\cF+\nu_\cE,\quad u =\frac{u'}{\norm{u'}}.
    \end{equation}
    The \emph{Householder rotation} $H(\nu_\cF,\nu_\cE)$ is defined as
    \begin{equation}\label{eq:H definition}
      H(\nu_\cF,\nu_\cE) = 2 u u\transpose-I.
    \end{equation}
  \end{definition}
  Here $H$ is a rotation mapping $\hat{\nu_\cF}$ to $\hat{\nu_\cE}$, as shown by the following.
  \begin{proposition}\label{prop:HProperty}
    The matrix $H$ has the following properties:
    \begin{enumerate}
    \item It is a rotation, i.e.
      \begin{enumerate}
      \item\label{it:orthonormality} $H\transpose H=I$;
      \item\label{it:determinant} $\det(H)=1$.
      \end{enumerate}
    \item\label{it:transformation} $\nu_\cE=H \nu_\cF$.
    \end{enumerate}
  \end{proposition}
  \begin{proof}
  	See Appendix \ref{proof:HProperty}.
  \end{proof}

@@

\subsubsection{Alternating Directions Method of Multipliers (ADMM)}\label{chapter:ADMM review}
The basic idea of the ADMM-based solver introduced in \cite{yang2020multi} is to separate the constraints from the objective function using a different set of variables $\vz$, and then solve separately in \eqref{eq:general-problem}.
More specifically, we rewrite the constraint $\vq\in\varOmega$ using an indicator function $\varTheta$, and include it in the objective function (details can be found in \cite{yang2020multi}). 
%In the traditional application of ADMM, the variables $\vz$ are simple duplicates of $\vx$ projected to the constraint set $\Omega$. However, in path planning problems, some constraints are non-convex, rendering the projection step more difficult (due to the presence of multiple local minima).
We allow $\vz = D(\vq)$ to replicate an arbitrary function of the main variables $\vq$ (instead of being an exact copy in general ADMM formulation), to transform constraint $\vq\in\Omega$ to $D(\vq) \in \sZ$ to allow for an easier projection step in \Cref{eq:z-update}. In summary, we transform \Cref{eq:general-problem} into
\begin{equation}\label{eq:ADMMSetConstraint_modified}
	\begin{aligned}
		&\max\quad \varPhi(\vq)+\varTheta(\vz) \\
		& \begin{array}{r@{\quad}c}
			s.t.& D(\vq)-\vz=0
		\end{array} 
	\end{aligned}
\end{equation}
where $D(\vq)= [D_1(\vq)^T,\dots,D_l(\vq)^T]^T$ is a vertical concatenation of different functions for different constraints. This makes each constraint set $\mathcal{Z}_i$ independent and thus can be computed separately in later updating steps. $D_i(\vq)$ is chosen that the new constraint set~$\cZ_i$ becomes simple to compute which is illustrated in later secitons. % and the function~$D_i(\vq)$ can be used to select only the subset of the variables on which a constraint depends, speeding up computations.

The update steps of the algorithm are then derived as \cite{Boyd2011}:
\begin{subequations}\label{eq:ADMMupdate}
	\begin{align}
		\label{eq:admm-mod-update-q}\vq^{k+1}&:=\argmin_\vq(\varPhi(\vq^k) +\frac{\rho}{2}\norm{D(\vq)-\vz^k+\vu^k}_2^2),\\
		\vz^{k+1}&:=\Pi_\mathcal{Z}(D(\vq^{k+1})+\vu^k), \label{eq:z-update} \\
		\vu^{k+1}&:=\vu^k+D(\vq^{k+1})-\vz^{k+1},
	\end{align}
\end{subequations}
where $\Pi_\mathcal{Z}$ is the new projection function to the modified constraint set $\mathcal{Z}$, $\vu$ represents a scaled dual variable that, intuitively, accumulates the sum of primal residuals
\begin{equation}\label{eq:primal-residual}
	\vr^{k}=D(\vq^{k+1})-\vz^{k+1}.
\end{equation}
Checking the primal residuals alongside with the dual residuals 
\begin{equation}\label{eq:dual-residual}
	\vs^{k}=-\rho(\vz^{k}-\vz^{k-1})
\end{equation}
after each iteration, the steps are reiterated until convergence when the primal and dual residuals are small, or divergence when primal and dual residual remains large after a fixed number of iterations.

We now provide the functions $D(\vq)$, the sets $\cZ$, and the corresponding projection operators $\Pi_\cZ$ for security constraints including co-observation security constraints (\Cref{sec:co-observation-constraint}), and reachability constraints (\Cref{sec:ellipsoid-point}-\Cref{eq:region_ellipsoid_constraint}). The latter are based on the definition of \emph{ellipse-region-constraint} (\Cref{sec:reachability}).
Formulation of traditional path planning constraints like velocity constraints, convex obstacle constraints can be found in \cite{yang2020multi}.
